---
title: "R Notebook"
output: 
  html_document:
    keep_md: yes
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    theme: cerulean #flatly #journal #cosmo
  pdf_document: default
---


# Preamble

## Call libraries
```{r}
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(ggthemes)
library(infer)
```


## Set up colour palette
```{r}
# The palette as per SF:
phs_palette <- c("#99DAF5", "#004785", "#C027B9", "#82BB25")

# # To use for fills, add
#   scale_fill_manual(values = phs_palette)
# 
# # To use for line and point colors, add
#   scale_colour_manual(values = phs_palette)
```


# Number of beds - Data Wrangling

## View bed data
```{r}
beds <- read_csv("data/beds_by_nhs_board_of_treatment_and_specialty.csv") 


glimpse(beds)
```

### Check of out of scope quarters
```{r}
beds %>% 
  distinct(Quarter)
```


### Check data  for NA values
```{r}
quarters_out_of_scope <- c("2016Q3", "2016Q4", "2017Q1", "2017Q2", "2017Q3",
                           "2017Q4", "2018Q1", "2018Q2", "2018Q3", "2018Q4", 
                           "2019Q1", "2019Q2", "2019Q3", "2019Q4")

beds %>%
  filter(Quarter %in% quarters_out_of_scope) %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```

### Keep data related to our time frame of interest 
```{r}
quarters_out_of_scope <- c("2016Q3", "2016Q4", "2017Q1", "2017Q2", "2017Q3",
                           "2017Q4", "2018Q1", "2018Q2", "2018Q3", "2018Q4", 
                           "2019Q1", "2019Q2", "2019Q3", "2019Q4")
```


```{r}
beds_oos <- beds %>%
  filter(Quarter %in% quarters_out_of_scope)
```



```{r}
glimpse(beds_oos)
```



### How many HB items do we have?
```{r}
beds_oos %>%
  distinct(HB)
```


### Keep SpecialtyName == All Acute for the analysis
```{r}
beds_oos <- beds_oos %>%
  filter(SpecialtyName == "All Acute")
```


## View health board data
```{r}
health_board_labels <- read_csv("data/hb14_hb19.csv") 

glimpse(health_board_labels)
```

### View all HB values
```{r}
health_board_labels %>% 
  distinct(HB, HBName, HBDateEnacted)
```



### Remove junk
```{r}
health_board_labels <- health_board_labels %>% 
  select(HB, HBName)
  

glimpse(health_board_labels)
```



## View special health board data

**Note:** I had to use the following arguments to read the csv file, because the "The Golden Jubilee National Hospital" in the SHName column had a special Unicode character that I couldn't get rid off (if I kept the unicode character, I couldn't ``glimpse`` the data).
```{r}
shb_labels <- read_csv("data/special-health-boards_19022021.csv", locale = locale(encoding = "ASCII")) 

glimpse(shb_labels)
```

### View all SHB values
```{r}
shb_labels %>% 
  distinct(SHB, SHBName)
```

### Remove junk
```{r}
shb_labels <- shb_labels %>% 
  select(SHB, SHBName)
  

glimpse(shb_labels)
```



### View country label data
```{r}
country_labels <- read_csv("data/geography_codes_and_labels_country.csv") 


glimpse(country_labels)
```


# Join tables 

## Join bed & special health board
```{r}
beds_oos <- beds_oos %>%
  left_join(shb_labels, by = c("HB" = "SHB"))

glimpse(beds_oos)
```

```{r}
beds_oos %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```



## Joint bed & health board
```{r}
beds_oos <- beds_oos %>%
  left_join(health_board_labels, "HB")

glimpse(beds_oos)
```

```{r}
beds_oos %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```



```{r}
# beds <- beds %>% 
#   mutate(Country = if_else(!is.na(Country.x), Country.x, Country.y)) %>% 
#   select(-Country.x, -Country.y)
# 
# glimpse(beds) 
```



## Join bed & country labels
```{r}
beds_oos <- beds_oos %>%
  left_join(country_labels, by = c("HB" = "Country"))

glimpse(beds_oos)
```



```{r}
beds_oos %>%
  # filter(Quarter %in% quarters_in_scope) %>%
  # filter(SpecialtyName == "All Acute") %>% 
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```

##  Tidy up finalised bed_oos table
```{r}
beds_oos <- beds_oos %>%
  clean_names()

glimpse(beds_oos)
```


**Sanity Check - There should be on average 13,270 available staffed beds in Scotland during quarter ending December 2019 (PHS report)**
```{r}
beds_oos %>%
  # filter(hb_name == "NHS Borders")
  filter(country_name == "Scotland")

```



# KPI plots (Scotland: 2016Q3 - 2019Q4)


## Selected Geography: Scotland
```{r}
beds_oos %>%
  filter(country_name == "Scotland") %>% 
  ggplot() +
  aes(x = quarter, y = total_occupied_beds, group = country_name, colour = country_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))


beds_oos %>%
  filter(country_name == "Scotland") %>%
  ggplot() +
  aes(x = quarter, y = average_available_staffed_beds, group = country_name, colour = country_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))


beds_oos %>%
  filter(country_name == "Scotland") %>%
  ggplot() +
  aes(x = quarter, y = average_occupied_beds, group = country_name, colour = country_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))

beds_oos %>%
  filter(country_name == "Scotland") %>%
  ggplot() +
  aes(x = quarter, y = percentage_occupancy, group = country_name, colour = country_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))
```

### Sanity check No. 1. 

Check if I pick in the location column the aggregated results of value S08000031 for NHS Greater Glasgow and S08000020 for Grampian
```{r}
beds_oos %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde") %>% 
  filter(hb_name == "NHS Grampian") %>% 
  filter(location == hb) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy))
```



```{r}
beds_oos %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde" ) %>% 
  filter(hb_name == "NHS Grampian" ) %>% 
  group_by(location) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy)) 
```

## Selected Geography: NHS Board
```{r}
beds_oos %>%
  filter(hb_name == "NHS Greater Glasgow and Clyde") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, y = total_occupied_beds, group = country_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))


beds_oos %>%
  filter(hb_name == "NHS Greater Glasgow and Clyde") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, y = average_available_staffed_beds, group = country_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))


beds_oos %>%
  filter(hb_name == "NHS Greater Glasgow and Clyde") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, y = average_occupied_beds, group = country_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))

beds_oos %>%
  filter(hb_name == "NHS Greater Glasgow and Clyde") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, y = percentage_occupancy, group = country_name, colour = hb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))
```


### Sanity check No. 2.
Check to see if I pick in the location column the aggregated results of value SB0801 
```{r}
beds_oos %>%
  # filter(hb_name == "NHS Greater Glasgow and Clyde" ) %>% 
  filter(shb_name == "The Golden Jubilee National Hospital" ) %>% 
  group_by(location) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy)) 
```

```{r}
beds_oos %>%
  filter(shb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>% 
  summarise(avg_tot_occ_beds = mean(total_occupied_beds),
            avg_avail_staffed_beds = mean(average_available_staffed_beds),
            avg_occupied_beds = mean(average_occupied_beds),
            avg_percentage_occupancy = mean(percentage_occupancy))
```



## Selected Geography: Special Health Board
```{r}
beds_oos %>%
  filter(shb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, y = total_occupied_beds, group = country_name, colour = shb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))


beds_oos %>%
  filter(shb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, y = average_available_staffed_beds, group = country_name, colour = shb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))


beds_oos %>%
  filter(shb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>% 
  ggplot() +
  aes(x = quarter, y = average_occupied_beds, group = country_name, colour = shb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))

beds_oos %>%
  filter(shb_name == "The Golden Jubilee National Hospital") %>%
  filter(location == hb) %>%  
  ggplot() +
  aes(x = quarter, y = percentage_occupancy, group = country_name, colour = shb_name) +
  geom_line() +
  geom_point() +
  scale_colour_manual(guide = "none", values = phs_palette[2]) +
  ylim(c(0, NA)) +
  theme(axis.text.x = element_text(angle = 30, vjust = +1, hjust = 1))
```


# Save cleaned data

Write updated beds_oos data to a CSV file.
```{r}
 write_csv(beds_oos, here("beds_oos_clean.csv"))
```



# Investigate distribution of NHS data for Scotland

## Histogram
```{r}
beds_oos %>%
  select(total_occupied_beds, country_name) %>% 
  filter(country_name == "Scotland") %>% 
  ggplot() +
  aes(x = total_occupied_beds) +
  geom_histogram(col = "white", binwidth = 10000)


beds_oos %>%
  select(average_available_staffed_beds, country_name) %>% 
  filter(country_name == "Scotland") %>% 
  ggplot() +
  aes(x = average_available_staffed_beds) +
  geom_histogram(col = "white", binwidth = 200)


beds_oos %>%
  select(average_occupied_beds, country_name) %>% 
  filter(country_name == "Scotland") %>% 
  ggplot() +
  aes(x = average_occupied_beds) +
  geom_histogram(col = "white", binwidth = 200)


beds_oos %>%
  select(percentage_occupancy, country_name) %>% 
  filter(country_name == "Scotland") %>% 
  ggplot() +
  aes(x = percentage_occupancy) +
  geom_histogram(col = "white", binwidth = 1)
```
## Box Plot
```{r}
beds_oos %>%
  select(total_occupied_beds, country_name) %>% 
  filter(country_name == "Scotland") %>% 
  ggplot() +
  aes(x = total_occupied_beds) +
  geom_boxplot()


beds_oos %>%
  select(average_available_staffed_beds, country_name) %>% 
  filter(country_name == "Scotland") %>% 
  ggplot() +
  aes(x = average_available_staffed_beds) +
  geom_boxplot()


beds_oos %>%
  select(average_occupied_beds, country_name) %>% 
  filter(country_name == "Scotland") %>% 
  ggplot() +
  aes(x = average_occupied_beds) +
  geom_boxplot()


beds_oos %>%
  select(percentage_occupancy, country_name) %>% 
  filter(country_name == "Scotland") %>% 
  ggplot() +
  aes(x = percentage_occupancy) +
  geom_boxplot()
```

## Boostrap Sampling distribution

```{r}
bootstrap_distn <- beds_oos %>%
  select(total_occupied_beds, country_name) %>% 
  filter(country_name == "Scotland") %>% 
  specify(response = total_occupied_beds) %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "mean")

bootstrap_distn %>%
  visualise(binwidth = 10000)
```

## Calculate a 95% CI for ``mean(total_occupied_beds)``, and visualise it on the distribution
```{r}
lot_area_ci999 <- bootstrap_distn %>%
  get_ci(level = 0.999, type = "percentile")
lot_area_ci999


bootstrap_distn %>%
  visualise(bins = 30) +
  shade_ci(endpoints = lot_area_ci999)
```



# Doodles

<!-- ```{r} -->
<!-- country_name_beds <- unique(beds$country_name) %>%  -->
<!--   discard(is.na) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- hb_name_beds <- unique(beds$hb_name) %>%  -->
<!--   discard(is.na) -->

<!-- ``` -->


<!-- ```{r} -->
<!-- shb_name_beds <- unique(beds$shb_name) %>%  -->
<!--   discard(is.na) -->

<!-- ``` -->





















